Jug: Software for Reproducible Computation

### 2012-02-23 paper.tex
TXT Better text

\citet{Fomel2007} proposed a system built on top of Scons which shared
superficially similar design. Scons, like Make, is a build system. It supports
spawning parallel jobs using multiple threads. In the original use case, the
tasks are delegated to other commands and the operating system can take care of
parallelism. If, however, the tasks are computationally intensive in Python,
contention for the global interpreter lock will limit the amount of real
parallelism.\footnote{In the most commonly used Python interpreters, there is a
lock that prevents more than one thread from simultaneously executing
interpreted code, although they can execute non-interpreted code, as calling an
external programme or certain external libraries.} Ruffus is another
Python-based solution which supports parallel execution of pipelines
\citep{Goodstadt16092010}. There is a strong dependency on the filesystem,
which makes its use very different from jug usage, which strives to make its
usage similar to straight Python usage. Using the Ruby programming language,
\citet{mishima2011} proposed Pwrake.

@@

Jug is designed around tasks. A task is defined as a Python function and a set
of arguments, which may be Python values or the output of other tasks. A task
should be a pure function, but there is no way of mandating this in the
language.

@@

This defines the task dependency structure represented in
Figure~\ref{fig:jug-deps}. As we can see, all of the \code{count} operations
can be run in parallel, while the \code{mean} operation must wait the result of
all of the other computation. The dependency structure is always a \textsc{dag}
(directed acyclic graph). Only by bypassing the normal use mechanisms is it
even possible to construct a cycle.

@@

@@

The output of computation can be obtained from jug in two ways: (1) One can
simply write a task that outputs the desired format. (2) They can be inspected
interactively using the \emph{shell} subcommand. It is expected that the first
option will be used for the final version of the computation, whilst the second
one is most helpful during development.

@@

Therefore, I chose to force the user to explicitly invalidate his results,
while supporting automatic dependency discovery to make it easier for the user.
The recommendation is still that the user run the full pipeline from start to
finish once they are satisfied with the state of the code and certainly before
publication.

@@

Jug has been available in the Python Package Index\footnote{The Python Package
Index, PyPI, is accessible at \url{http://pypi.python.org}.} since May~2009
and has been downloaded over 8000~times (some of these downloads may, of
course, represent upgrades, and users which downloaded it from some other
source will not be counted).% FIXME: Check download number. At 8112 12/8/2011

>>>>

\citet{Fomel2007} proposed a system built on top of Scons which shared
superficially similar design. Scons, like Make, is a build system. It supports
spawning parallel jobs using multiple threads. In the original use case, the
tasks are delegated to other commands and the operating system can take care of
parallelism. If, however, the tasks are computationally intensive in Python,
contention for the global interpreter lock will limit the amount of real
parallelism.\footnote{In the most commonly used Python interpreters, there is a
lock that prevents more than one thread from simultaneously executing
interpreted code, although they can execute non-interpreted code, as calling an
external programme or certain external libraries.} Ruffus is another
Python-based solution which supports parallel execution of pipelines
\citep{Goodstadt16092010}. There is a strong dependency on the filesystem,
which makes its use very different from jug usage, which strives to make its
usage similar to straight Python usage. Using the Ruby programming language,
\citet{mishima2011} proposed Pwrake, which supports parallel execution of
bioinformatics workflows written (or accessible) in that language.

@@

Jug is designed around tasks. A task is defined as a Python function and a set
of arguments, which may be Python values or the output of other tasks.

@@

This code defines the task dependency structure represented in
Figure~\ref{fig:jug-deps}. As we can see, all of the \code{count} operations
can be run in parallel, while the \code{mean} operation must wait the result of
all of the other computation. The dependency structure is always a \textsc{dag}
(directed acyclic graph).

@@

This makes jug usable in batch-based computer cluster environments, which are
quite common in research institutions.

@@

The output of computation can be obtained from jug in two ways: (1) One can
simply write a task that writes to a file in the desired format. (2) They can
be inspected interactively using the \emph{shell} subcommand. It is expected
that the first option will be used for the final version of the computation,
whilst the second one is most helpful during development.

@@

Therefore, I chose to force the user to explicitly invalidate his results,
while supporting automatic dependency discovery to make it easier for the user.
The recommendation is still that the user run the full pipeline from start to
finish once they are satisfied with the state of the code and certainly before
publication, but the pipeline development stage can be more agile.

@@

Jug has been available in the Python Package Index\footnote{The Python Package
Index, PyPI, is accessible at \url{http://pypi.python.org}.} since May~2009
through February~2012 and has been downloaded over 10,000~times (some of these
downloads may, of course, represent upgrades, and users which downloaded it
from some other source will not be counted).
